---
title: "Interdependence between Short-Term and Long-Term Treasury Yields"
author: "Aayush Khandekar"
date: '`December 2nd, 2024'
output:
  html_document: default
  pdf_document: default
---

```{r}

library(forecast)
library(tidyverse)
library(vars)
library(ggplot2)

```

```{r}

data <- read.csv("data.csv")

# Differencing
data$c1_t <- c(NA, diff(data$DGS1))
data$c10_t <- c(NA, diff(data$DGS10))

data <- na.omit(data)

# Fit AR models for c1_t
ar_c1 <- auto.arima(data$c1_t, max.p = 5, max.q = 0, stationary = TRUE)
summary(ar_c1)

# Fit AR models for c10_t
ar_c10 <- auto.arima(data$c10_t, max.p = 5, max.q = 0, stationary = TRUE)
summary(ar_c10)

```
```{r}

ggplot(data, aes(x = 1:nrow(data), y = c1_t)) +
  geom_line(color = "black") +
  labs(title = "1-Year Treasury Yield Changes Over Time",
       x = "Time", y = "Change in 1-Year Yield")

ggplot(data, aes(x = 1:nrow(data), y = c10_t)) +
  geom_line(color = "black") +
  labs(title = "10-Year Treasury Yield Changes Over Time",
       x = "Time", y = "Change in 10-Year Yield")

```

```{r}

# Bivariate autoregressive model for two change series
change_series <- data.frame(c1_t = data$c1_t, c10_t = data$c10_t)

# Fit a VAR model
var_model <- VAR(change_series, p = 2, type = "const")  
summary(var_model)

```

```{r}

irf_result <- irf(var_model, impulse = "c1_t", response = "c10_t", n.ahead = 20, boot = TRUE)

plot(irf_result, main = "Impulse Response: c1_t Shock on c10_t")

```
The impulse response analysis reveals that a shock to the 1-year yield causes an immediate and significant rise in the 10-year yield, with the impact peaking at around 0.20 and fading within 5 periods. 

This provides clear visual and statistical evidence that short-term rate movements influence long-term rates, supporting the transmission of monetary policy through the yield curve.


```{r}

Bmat <- matrix(c(NA, 0, NA, NA), nrow = 2, byrow = TRUE)

svar_model <- SVAR(var_model, Bmat = Bmat, estmethod = "scoring")
summary(svar_model)

```


```{r}

acf(data$c1_t, main = "ACF of 1-Year Yield Change")
pacf(data$c1_t, main = "PACF of 1-Year Yield Change")

```
```{r}

fitted_vals <- fitted(ar_c1)

plot(data$c1_t, type = "l", col = "black", lwd = 1.5, ylab = "Yield Change", main = "Actual vs Fitted - AR(2) Model")
lines(fitted_vals, col = "red", lwd = 2)
legend("topright", legend = c("Actual", "Fitted"), col = c("black", "red"), lty = 1)

```


```{r}

# Lagged short-term change
data$c1_t_lag1 <- dplyr::lag(data$c1_t, 1)

# Scatter plot
library(ggplot2)
ggplot(data, aes(x = c1_t_lag1, y = c10_t)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", col = "red") +
  labs(title = "Lagged 1Y Yield Change vs Current 10Y Yield Change",
       x = "Lag 1 of 1-Year Yield Change",
       y = "10-Year Yield Change")

```

The scatter plot demonstrates a positive and statistically significant relationship between the previous day's change in the 1-year Treasury yield and the current day's change in the 10-year yield. This supports the hypothesis that short-term interest rate movements influence long-term rates with a lag, consistent with the expectations theory and results from the VAR model.



The VAR model reveals significant interdependencies between $c^1_t$ and $c^{10}_t$.

Specifically, $c^{10}_t$ at lag 1 significantly influences $c^1_t$ ($0.237$, $p = 0.004$), while $c^1_t$ has a negligible effect on $c^{10}_t$ ($-0.009$, $p = 0.817$).

This suggests that long-term rates have a stronger impact on short-term rates, reflecting their persistence over time.

In contrast, the single AR models primarily capture the autocorrelation within each series, such as the significant AR(2) terms for $c^1_t$ ($ar1 = 0.451$, $ar2 = -0.248$) and $c^{10}_t$ ($ar1 = 0.383$, $ar2 = -0.233$, $ar3 = 0.070$), but they miss the cross-series dynamics evident in the VAR model.
